---
title: "class19"
author: 'Jibin (PID: A53300326)'
date: "2021/12/1"
output: pdf_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Section 1. Proportion of G/G in a population

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
head(mxl)
```

```{r}
table(mxl$Genotype..forward.strand.)/nrow(mxl)*100
```

The proportion is 14.0625%

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
```

```{r}
table(gbr$Genotype..forward.strand.)/nrow(gbr)*100
```


This variant that is associated with childhood asthma is more frequent in the GBR population than MKL population.

# Section 4: Population Scale Analysis

> Q13: Read this file into R and determine the sample size for each genotype and their corresponding median expression levels for each of these genotypes.

```{r}
expr <- read.table("rs8067378_ENSG00000172057.6.txt")
head (expr)
```

```{r}
summary(expr)
table(expr$geno)
```

```{r}
library(ggplot2)
ggplot(expr)+ aes(geno, exp, fill=geno)+
  geom_boxplot(notch=T)+
  geom_point(position=position_jitter(h=0.15,w=0.2), alpha=0.2, size=2)+
  labs(x="Genotype", y= "Expression")
```

> Q14: Generate a boxplot with a box per genotype, what could you infer from the relative expression value between A/A and G/G displayed in this plot? Does the SNP effect the expression of ORMDL3

Yes. The G/G SNP dose show a lower expression of ORMDL3 than A/A phenotype.

