---
title: "class15"
author: 'Jibin (PID: A53300326)'
date: "2021/11/17"
output: html_document
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

## load the contData nd colData

We need 2 things
- countData
- colData


```{r}
library(BiocManager)
library(DESeq2)
```



```{r}
counts <- read.csv("airway_scaledcounts.csv", row.names = 1)
metadata <- read.csv("airway_metadata.csv")
```

```{r}
head(counts)
```

```{r}
head(metadata)
```


Side-note:
Let's check the corespondence of the metadata and count data setup.

```{r}
metadata$id
```

```{r}
colnames(counts)
```

We can use the `==` thing to see if they are the same

```{r}
metadata$id == colnames(counts)
```

```{r}
all(c(T,T,T,T,T,T,F))
all(metadata$id == colnames(counts))
```

## Compare control to treated

First we need to access all the control columns in our counts data.

```{r}
control.inds <- metadata$dex == "control"
metadata[control.inds, ]$id
```

Use these ids to access just the control columns of our counts data

```{r}
head(counts[, control.inds])
```

```{r}
control.mean <- rowMeans(counts[, control.inds])
head(control.mean)
```

Do the same for drug treated

```{r}
treated.inds <- metadata$dex == "treated"
metadata[treated.inds, ]$id
```


```{r}
treated.mean <- rowMeans(counts[, treated.inds])
head(treated.mean)
```

we will combine our means count data for bookkeeping purposes


```{r}
meancounts<-  data.frame (control.mean, treated.mean)
```

There are `r nrow(counts)`in this dataset

```{r}
nrow(counts)
```

## Compare the control and treated


```{r}
plot(meancounts)
```

This would benefit from a long transform! Let's plot on a log scale

```{r}
plot(meancounts, log="xy")
```
We often use log trasforamtions as they make life much nicer in this world...

```{r}
log2(40/20)
```

Cool. I like log2!

```{r}
meancounts$log2fc <- log2(meancounts[,"treated.mean"]/meancounts[,"control.mean"])
head(meancounts)
```

The `which ()` function tells us the indices of TRUE netries in a logical vector.

```{r}
which (c(T, F, T))
```


```{r}
zero.vals <- which(meancounts[,1:2]==0, arr.ind=TRUE)

to.rm <- unique(zero.vals[,1])
mycounts <- meancounts[-to.rm,]
head(mycounts)
nrow(mycounts)
```

```{r}
up.ind <- mycounts$log2fc > 2
sum(up.ind)
down.ind <- mycounts$log2fc < (-2)
sum(down.ind)
```

What the percentage is this?

```{r}
round(sum(mycounts$log2fc > 2)/nrow(mycounts)*100, 2)
```

# DESeq2 analysis

```{r}
library(DESeq2)
citation("DESeq2")
```


```{r}
dds <- DESeqDataSetFromMatrix(countData=counts, 
                              colData=metadata, 
                              design=~dex)
dds
```

```{r}
dds <- DESeq(dds)
```


```{r}
res <- results(dds)
head (res)
```
We can summarize some basic tallies using the summary function.

```{r}
summary(res)
```
If the adjusted p value cutoff will be a value other than 0.1, alpha should be set to that value:


```{r}
res05 <- results(dds, alpha=0.05)
summary(res05)
```


# A volcano plot

this is a very common data viz of this

```{r}
plot( res$log2FoldChange,  -log(res$padj), 
      xlab="Log2(FoldChange)",
      ylab="-Log(P-value)")
```

```{r}
plot( res$log2FoldChange,  -log(res$padj), 
 ylab="-Log(P-value)", xlab="Log2(FoldChange)")

# Add some cut-off lines
abline(v=c(-2,2), col="darkgray", lty=2)
abline(h=-log(0.05), col="darkgray", lty=2)
```

```{r}
# Setup our custom point color vector 
mycols <- rep("gray", nrow(res))
mycols[ abs(res$log2FoldChange) > 2 ]  <- "red" 

inds <- (res$padj < 0.01) & (abs(res$log2FoldChange) > 2 )
mycols[ inds ] <- "blue"

# Volcano plot with custom colors 
plot( res$log2FoldChange,  -log(res$padj), 
 col=mycols, ylab="-Log(P-value)", xlab="Log2(FoldChange)" )

# Cut-off lines
abline(v=c(-2,2), col="gray", lty=2)
abline(h=-log(0.1), col="gray", lty=2)
```

```{r}
library(EnhancedVolcano)

```

## Adding annotation data

We want to add meaningful gene names to our dataset so we can ake some sense of what is going on here 

For this will will use two bioconductor packages, one dose the work and is called **AnnotationDbi** and  the other contains the data we are going to map between and is called **org.Hs.eg.db**


```{r}
library("AnnotationDbi")
library("org.Hs.eg.db")
```


```{r}
columns(org.Hs.eg.db)
```

```{r}
res$symbol <- mapIds(org.Hs.eg.db,
                     keys=row.names(res), # Our genenames
                     keytype="ENSEMBL",        # The format of our genenames
                     column="SYMBOL",          # The new format we want to add
                     multiVals="first")
```

```{r}
head (res)
```

```{r}
ord <- order( res$padj )
#View(res[ord,])
head(res[ord,])
```


```{r}
library(EnhancedVolcano)
x <- as.data.frame(res)

EnhancedVolcano(x,
    lab = x$symbol,
    x = 'log2FoldChange',
    y = 'pvalue')
```

# Let's finally save our results to data

```{r}
write.csv(res[ord,], "deseq_results.csv")
```

#Pathway Analysis

Let's try to bring some biology insights back into this work

```{r paged.print=FALSE}
library(pathview)
library(gage)
library(gageData)

data(kegg.sets.hs)

# Examine the first 2 pathways in this kegg set for humans
head(kegg.sets.hs, 2)
```

Before we can useKEGG we need to get oiur gene identifiers in the correct format for KEGG, which is ENTREZ format in this case.


```{r}
columns(org.Hs.eg.db)
```


```{r}
res$entrez <- mapIds(org.Hs.eg.db,
                     keys=row.names(res),
                     column="ENTREZID",
                     keytype="ENSEMBL",
                     multiVals="first")

res$uniprot <- mapIds(org.Hs.eg.db,
                     keys=row.names(res),
                     column="UNIPROT",
                     keytype="ENSEMBL",
                     multiVals="first")

res$genename <- mapIds(org.Hs.eg.db,
                     keys=row.names(res),
                     column="GENENAME",
                     keytype="ENSEMBL",
                     multiVals="first")

head(res)
```

Assign names to this vector that are the gene IDs that KEGG wants.
```{r}
foldchanges = res$log2FoldChange
names(foldchanges) = res$entrez
head(foldchanges)
```

```{r}
# Get the results
keggres = gage(foldchanges, gsets=kegg.sets.hs)
```

We can look at the attributes() of this or indeed any R object.

```{r}
attributes(keggres)
```

```{r}
# Look at the first three down (less) pathways
head(keggres$less, 3)
```

```{r}
pathview(gene.data=foldchanges, pathway.id="hsa05310")
```

! [](hsa05310.pathview.png)


# Plotting counts for genes of interest
```{r}
i <- grep("CRISPLD2", res$symbol)
res[i,]
```

```{r}
plotCounts(dds, gene="ENSG00000103196", intgroup="dex")
```

```{r}
d <- plotCounts(dds, gene="ENSG00000103196", intgroup="dex", returnData=TRUE)
head(d)
```

```{r}
library(ggplot2)
ggplot(d, aes(dex, count, fill=dex)) + 
  geom_boxplot() + 
  scale_y_log10() + 
  ggtitle("CRISPLD2")
```


