---
title: "Class10_HalloweenMiniProject"
author: 'Jibin (PID: A53300326)'
date: "2021/10/29"
output: 
  pdf_document:
    latex_engine: xelatex
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```

# Class 10: Halloween Mini-Project

## 1. Importing candy data

```{r}
candy_file <- "candy-data.csv"

candy = read.csv(candy_file, row.names=1)
head(candy)
```

> Q1. How many different candy types are in this dataset?
> Q2. How many fruity candy types are in the dataset?

```{r}
nrow(candy)
sum(candy$fruity)
```

85 candy types and 38 fruity candy types.


## 2. What is your favorate candy?


> Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

3 Musketeers

> Q4. What is the winpercent value for “Kit Kat”?

```{r}
candy["Kit Kat", ]$winpercent
```

> Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

```{r}
candy["Tootsie Roll Snack Bars", ]$winpercent
```

```{r}
library("skimr")
skim(candy)
```

> Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

Winpercent is on a different sacle to others.

> Q7. What do you think a zero and one represent for the candy$chocolate column?

Zero indicates it's not chocolate while one means it's chocolate candy


> Q8. Plot a histogram of winpercent values

```{r}
hist(candy$winpercent)
```


> Q9. Is the distribution of winpercent values symmetrical?

No

> Q10. Is the center of the distribution above or below 50%?

Above 50%.

> Q11. On average is chocolate candy higher or lower ranked than fruit candy?

```{r}
chocolate <- candy$winpercent[as.logical(candy$chocolate)]
mean(chocolate)
```

```{r}
chocolate <- candy[as.logical(candy$chocolate), ]$winpercent
mean(chocolate)
fruit <- candy[as.logical(candy$fruit), ]$winpercent
mean(fruit)
```


> Q12. Is this difference statistically significant?

```{r}
t.test(chocolate, fruit)
```

Yes, the difference is significant.



3. Overall Candy Rankings

> Q13. What are the five least liked candy types in this set?

```{r}
head(candy[order(candy$winpercent),], n=5)
```


> Q14. What are the top 5 all time favorite candy types out of this set?

```{r}
tail(candy[order(candy$winpercent),], n=5)
```

> Q15. Make a first barplot of candy ranking based on winpercent values.

```{r}
library(ggplot2)

ggplot(candy) +
  aes(winpercent, rownames(candy))+
  geom_col()
```
> Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?

```{r}
library(ggplot2)

ggplot(candy) +
  aes(winpercent, reorder(rownames(candy), winpercent))+
  geom_col()
```

## Time to add some useful color

```{r}
my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
library(ggplot2)

ggplot(candy) +
  aes(winpercent, reorder(rownames(candy), winpercent))+
  geom_col(fill= my_cols)
```

Now, for the first time, using this plot we can answer questions like:
> Q17. What is the worst ranked chocolate candy?

Reeses Peanut Buttercup

> Q18. What is the best ranked fruity candy?

Starburst

4. Taking a look at pricepercent

```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

```{r}
library(ggrepel)

# How about a plot of price vs win
rownames(candy)= gsub("脮" , "'", row.names(candy))
my_cols <- rep(c("red", "black"), round(nrow(candy)/2))
my_cols = c(my_cols, "red")
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols , size=3.3, max.overlaps = 7)
```

> Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

It's Reese's Miniatures.


> Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

It's Hershey's Krackel.

```{r}
rownames(candy)= gsub("脮", "'", row.names(candy))
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```



5 Exploring the correlation structure

```{r}
library(corrplot)
cij <- cor(candy)
corrplot(cij)
```
> Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

fruity and chocolate.

> Q23. Similarly, what two variables are most positively correlated

chocolate and bar!

6. Principal Component Analysis

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

## Now we can plot our main PCA score plot of PC1 vs PC2.

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

# Make a new data-frame with our PCA results and candy data

```{r}
my_data <- cbind(candy, pca$x[,1:3])

p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")
```

```{r}
# library(plotly)
# ggplotly(p)
```

> Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

"Fruity"! Yes, that makes sense since it never taste both like candy and chocolate:)

```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```


