---
title: "Class8"
author: 'Jibin (PID: A53300326)'
date: "2021/10/22"
output: pdf_document
---
# kmeans

```{r}
tmp = c(rnorm(30,3), rnorm(30, -3))
data <-  cbind(x=tmp, y=rev(tmp))
hist(data)
plot(data)
```

Run `kmeans()` set k to 2 nstart 20. The thing with kmeans is you have to tell it how many clusters you want.

```{r}
km <- kmeans(data, centers=2, nstart =20)
km
```

>Q. How many points are in each clusters?

```{r}
km$size
```
>Q. What "component" of your result object details cluster assigment/membership?

```{r}
km$cluster
```


>Q. What "component" of your result object details cluster center?

```{r}
km$centers
```


>Q. Plot x colored by the kmeans cluster assigment and add cluster centers as blue points?

```{r}
plot(data, col=km$cluster)
points(km$centers, col="blue", pch=15, cex=2)
```

# Hierarchical Clustering

We will use the hclust() function on the same data as before and see how this method works.

```{r}
hc <- hclust(dist(data))
hc
```

```{r}
plot(hc)
abline(hc, h=7, col="red")
```

To find our membership vector we need to "cut" the tree and for this we use the `cutree()` function and tell it the height to cut at.


```{r}
plot(hc)
cutree(hc, h=7)
```

```{r}
grps <- cutree(hc, k=2)
```

```{r}
plot(data, col=grps)
```

```{r}
plot(hclust(dist(data), method="complete"))
plot(hclust(dist(data), method="average"))
plot(hclust(dist(data), method="single"))
```


# Principal Component Analysis (PCA)

PCA is a super useful analysis method when you have lots of dimensions in your data

# PCA of the UK_food

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url)
```

how many rows and columns are in x?

```{r}
dim(x)
```


```{r}
x
```

```{r}
rownames(x)<- x[,1]
x <- x[,1]
x
```

Q2. Which approach to solving the ‘row-names problem’ mentioned above do you prefer and why? Is one approach more robust than another under certain circumstances?

```{r}
url <- "https://tinyurl.com/UK-foods"
x <- read.csv(url, row.names = 1)
x
```

Q3: Changing what optional argument in the above barplot() function results in the following plot?

```{r}
barplot(as.matrix(x), col=rainbow(17),)
barplot(as.matrix(x), col=rainbow(17), beside=T)
```

Q5: Generating all pairwise plots may help somewhat. Can you make sense of the following code and resulting figure? What does it mean if a given point lies on the diagonal for a given plot?

```{r}
mycol <- rainbow(nrow(x))
pairs(x, col=mycol, pch=16)
```

## PCA to the rescue

Here we will use the base R function for PCA, which is called `prcomp()`. This function wants the transpose of the data.

```{r}
pca <- prcomp(t(x))
summary(pca)
```

```{r}
plot(pca)
```

We want score plot (a.k.a, PCA plot). Basically of PC1 vs PC2

```{r}
attributes(pca)
```

Q7. Complete the code below to generate a plot of PC1 vs PC2. The second line adds text labels over the data points

```{r}
plot(pca$x[,1:2])
text(pca$x[,1:2], labels = colnames(x))
```

Q8. Customize your plot so that the colors of the country names match the colors in our UK and Ireland map and table at start of this document.

```{r}
plot(pca$x[,1:2])
text(pca$x[,1:2], labels = colnames(x), col=c("orange", "pink", "light blue", "green"))
```


We can also examine the PCA "loading", which tell us how much the roifinal variable contribute to teach new PC...

```{r}
pca$rotation
barplot(pca$rotation[,1], las=2)
```
Q9: Generate a similar ‘loadings plot’ for PC2. What two food groups feature prominantely and what does PC2 maninly tell us about?

```{r}
pca$rotation
barplot(pca$rotation[,2], las=2)
```

## One more PCA for today

```{r}
url2 <- "https://tinyurl.com/expression-CSV"
rna.data <- read.csv(url2, row.names=1)
head(rna.data)
```

```{r}
dim(rna.data)
```

```{r}
pca.rna <- prcomp(t(rna.data), scale=TRUE)
summary(pca.rna)
```

```{r}
plot(pca.rna)
```

```{r}
plot(pca.rna$x[,1:2])
text(pca.rna$x[,1:2], labels = colnames(rna.data))
```

