---
title: "Class17"
author: 'Jibin (PID: A53300326)'
date: "2021/11/24"
output:
 pdf_document:
  latex_engine: xelatex
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```


# Import vaccination data

```{r}
vax <- read.csv( "covid19vaccinesbyzipcode_test.csv" )
head(vax)
```

> How many entries do have have?

```{r}
nrow(vax)
```

We can use the **skimr**package and the `skim()` function to get a quick overview of the structure of this dataset.

```{r}
skimr::skim(vax) 
```

Notice that one of these column is a date column. Working with time and dates get's annoying quickly. We can use the **lubridate** package to make this easy...

```{r}
library("lubridate")

today()
```

> Q. How many days since the first enty in the dataset?

```{r}
vax$as_of_date[1]
```


This will not work because out data column was read as character..
```{r}
#today() - vax$as_of_date[1]
```

```{r}
d <- ymd(vax$as_of_date)
```

```{r}
today() - d[1]
```

I will make the `as of date` column Data format

```{r}
vax$as_of_date <- ymd(vax$as_of_date)
```

> Q. When was the dataset last updated? What is the last date in this dataset? How many days since the last update?


```{r}
today() - vax$as_of_date[nrow(vax)]
```
> Q. How many days dose the dataset span?


```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```

> Q. How many different ZIP code areas are?

```{r}
AZIP <- unique(vax$zip_code_tabulation_area)
length(AZIP)
```

To work with ZIP codes we can use **zipcodeR**

```{r}
library("zipcodeR")
```

```{r}
geocode_zip('92037')
```

```{r}
zip_distance('92037','92109')
```

```{r}
reverse_zipcode(c('92037', "92109") )
```

## Focus in on San Diego County

We want to subset the full CA `vax` data to down to just San Diego County.

```{r}
inds <- vax$county == "San Diego"
nrow(vax[inds,])
```

Subsetting can get tedious and complicated quickly when you have multiple things want to subset by.

```{r}
library("dplyr")
```


We will use the `filter()` function to do our subsetting from now on.

```{r}
sd <- filter(vax, county == "San Diego")
nrow(sd)
```

More complicated subsetting...

```{r}
sd.20 <- filter(vax, county=="San Diego",
       age5_plus_population>20000)

nrow(sd.20)
```

> Q. What is the averaege vaccination rate of San Diegocount as of yesterday?

```{r}
sd.now <- filter(vax, county=="San Diego",
                 as_of_date=="2021-11-23")

head(sd.now)
```


```{r}
summary(sd.now$percent_of_population_fully_vaccinated)
```

> Q. Make a histogram of these values

```{r}
hist(sd.now$percent_of_population_fully_vaccinated)
```

This is going to be susceptible to being skewed by ZIP code areas with small populations. This will have big effects for just a small number of unvax-ed folks.

> Q. What is the population of the 92037 zip ode area?

```{r}
ucsd <- filter(sd.now, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```


> Q. What is the average vaccination value for this UCSD/La Jolla ZIP code area?

```{r}
ucsd$percent_of_population_fully_vaccinated
```


```{r}
LJ2 <- filter(sd.now, zip_code_tabulation_area=="92122")
LJ2$percent_of_population_fully_vaccinated
```

Time series of vaccination rate for a given ZIP code area.

Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area

```{r}
Lj <- filter(vax, zip_code_tabulation_area=="92037")
```


```{r}
library(ggplot2)
ggplot(Lj) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated) +
  geom_point() +
  geom_line(group=1) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent  Vaccinated")
```

Let's make this plot for all San Diego County ZIP code areas that have a population as least as large as 92037.

```{r}
sd.36 <- filter(vax, county=="San Diego",
             age5_plus_population>36144)
head(sd.36)
```

> How many ZIP code areas in San Diego county have a population larger than 92037?

```{r}
length(unique(sd.36$zip_code_tabulation_area))
```


Let's make the plot

```{r}
library(ggplot2)
ggplot(sd.36) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, group=zip_code_tabulation_area) +
  geom_line() +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent  Vaccinated")
```

> Q. Make a plot like this for the all ZIP code areas in the State with a population at least as larlge as LJ.

# Subset to all CA areas with a population as large as 92037

```{r}
ca <- filter(vax, age5_plus_population > 36144)
```

```{r}
length(unique(ca$zip_code_tabulation_area))
```


```{r}
library(ggplot2)
ggplot(ca) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, group=zip_code_tabulation_area) +
  geom_line(alpha=0.2) +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent  Vaccinated")
```

>Q. What is the mean across the state for these 36k+ population area?

```{r}
ca.now <- filter(ca, as_of_date=="2021-11-23")
summary(ca.now$percent_of_population_fully_vaccinated)
```


```{r}
ggplot(ca) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color="blue") +
  ylim(c(0,1)) +
  labs(x="Date", y="Percent  Vaccinated",
       title="Vaccination Rate Across California",
       subtitle="Only areas with a population above 36k are shown") +
  geom_hline(yintercept = 0.75, color="red")
```

