---
title: 'Class 6: R Functions'
author: 'Jibin (PID: A53300326)'
date: "2021/10/15"
output: pdf_document
---

## Quick Markdown intro

We can write text of course just like any file We can **style text to be bold** or **italic**

Do

- this
- and that
- and another thing

This is more text
and this is a new line

---
We can include some code:

```{r}
plot(1:10)
```

```{r}
plot(cars)
```
## Time to write a function

> **Q1**. Write a function grade() to determine an overall grade from a vector of student homework
assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an
NA value) this can be used as a score to be potentially dropped. Your final function should be
adquately explained with code comments and be able to work on an example class gradebook
such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]



```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```

First I want to find the lowest score. I can use the **min()** to ding it and the *which.min()* function to find where it is (i.e. its position in the vector)


```{r}
which.min(student1)
range(student1)
```

I can use minus to get everything in the vector but the lowest score.
```{r}
student1[-which.min(student1)]
```

Now I can call the **mean()** function to ge the average.

```{r}
mean(student1[-which.min(student1)])
```

```{r}
mean(student2, na.rm=TRUE)
```

```{r}
is.na(student2)
which(is.na(student2))
sum(is.na(student2))
student2[is.na(student2)] <- 0
```

```{r}
x <- student2
x[is.na(x)]= 0
x
mean(x[-which.min(x)])
```


```{r}
x <- student3
x[is.na(x)]= 0
x[ -which.min(x)]
mean(x[ -which.min(x)])
```

```{r}
student4 <- c(100, NA, 90, "90", 90, 90, 97, 80)
```

```{r}
grade <- function(x) {
  x <- as.numeric(x)
  x[is.na(x)]= 0
  mean(x[ -which.min(x)])
}
```

```{r}
grade2 <- function(x, drop.lowest=TRUE){
  x <- as.numeric(x)
  x[is.na(x)]=0
  if(drop.lowest){
    mean(x[ -which.min(x)])
    } else{mean(x)}
}
```

```{r}
grade2(student2, drop.lowest=FALSE)
```



And test if it works on a single vector

```{r}
grade(student1)
```

```{r}
gradebook <- "https://tinyurl.com/gradeinput"
scores <- read.csv(gradebook, row.names = 1)
scores
```

```{r}
apply(scores, 1, grade)
ans <-apply(scores, 1, grade)
```

> **Q2** Who is the top scoring student?

```{r}
which.max(ans)
```

>**Q3** From your analysis of the gradebook, which homework was toughest on students (i.e. obtained the lowest scores overall? [2pts]

```{r}
apply(scores, 2, mean, na.rm=TRUE)
```

##Replace or mask NA values to zero

```{r}
mask <- scores
  is.na(scores)
  mask[is.na(scores)]=0
  mask
apply(mask, 2, mean)
which.min(apply(mask, 2, mean))
```
> "Q4" Optional Extension: From your analysis of the gradebook, which homework was most
predictive of overall score (i.e. highest correlation with average grade score)? [1pt]


```{r}
cor(mask$hw4, ans)
apply(mask, 2, cor, ans)
```

```{r}
boxplot(scores)
```


```{r}
df <- data.frame(a=1:10, b=seq(200,400,length=10),c=11:20,d=NA)

rescale <- function(x, na.rm= TRUE){
  rng <- range(x, na.rm= na.rm)
  (x-rng [1])/(rng[2]-rng[1])}
```


## install.packages("bio3d")
```{r}
library(bio3d)
s1 <- read.pdb("4AKE") # kinase with drug
s2 <- read.pdb("1AKE") # kinase no drug
s3 <- read.pdb("1E4Y") # kinase with drug
s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor") 
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor") 
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")

hc <- hclust( dist( rbind(s1.b, s2.b, s3.b) ) )
plot(hc)
```

